---
title: 'Tables & Figures'
output:
  pdf_document: default
---

```{r setup, include = F}
options(knitr.kable.NA = '')
knitr::opts_chunk$set(echo = F, message = F, warning = F)
# webshot::install_phantomjs()

library(tidyverse)
library(gridExtra)
library(kableExtra)
library(scales)
library(usmap)

data_dir <- file.path('..', '..', 'data')

load(file.path(data_dir, 'tbl_fig_data_final.RData'))

theme_set(
  theme(
    text = element_text(size = 7),
    panel.background = element_blank(),
    plot.title = element_text(color = '#444444', size = 7, hjust = 0.5, face = 'bold'),
    axis.ticks = element_blank(),
    axis.title = element_text(face = 'bold'),
    legend.title = element_text(face = 'bold'),
    legend.key.size = unit(0.3, 'cm'),
    strip.text.x = element_text(size = 7, face = 'bold')
  )
)

color_palette <- c('#bbcfd7', '#d2c8bc', '#ba9a88')
```


## Figure 7: Orders and prospects purchased by research vs. ma/doctoral

```{r, fig.height = 3}
orders_prospects_purchased %>% 
  ggplot(aes(x = reorder(univ_id, -total_students), y = total_students, fill = univ_label)) +
  geom_bar(stat = 'identity', alpha = 0.6) +
  geom_text(aes(label = str_c(total_orders, if_else(total_orders == 1, ' order', ' orders')), y = total_students + 2000), vjust = 0, size = 1.8) +
  scale_y_continuous(labels = function(n) if_else(n < 1000000, str_c(n / 1000, 'K'), str_c(n / 1000000, 'M'))) +
  scale_fill_manual(values = color_palette, name = 'University type') +
  xlab('') + ylab('Number of prospects') +
  theme(
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()
  )
```


## Figure 8: Filters used in order purchases by research vs. ma/doctoral

```{r, fig.height = 4}
orders_filters %>% 
  add_row(univ_label = 'Research', filters = 'academic', num = 0) %>% 
  add_row(univ_label = 'MA/doctoral', filters = 'academic', num = 0) %>% 
  add_row(univ_label = 'Research', filters = 'geographic', num = 0) %>% 
  add_row(univ_label = 'MA/doctoral', filters = 'geographic', num = 0) %>% 
  add_row(univ_label = 'Research', filters = 'demographic', num = 0) %>% 
  add_row(univ_label = 'MA/doctoral', filters = 'demographic', num = 0) %>% 
  mutate(
    filters_label = recode_factor(
      filters,
      'gender' = 'Gender',
      'race' = 'Race',
      'demographic' = '  ',
      'intl' = 'International',
      'cbsa' = 'CBSA',
      'segment' = 'Segment',
      'geomarket' = 'Geomarket',
      'zip' = 'Zip code',
      'states_fil' = 'State',
      'geographic' = ' ',
      'ap_score' = 'AP score',
      'rank' = 'Rank',
      'psat' = 'PSAT',
      'sat' = 'SAT',
      'gpa' = 'GPA',
      'academic' = '',
      'hsgrad_class' = 'HS grad class'
    )
  ) %>% 
  ggplot(aes(x = filters_label, y = num)) +
  geom_bar(stat = 'identity', fill = color_palette[[2]]) +
  geom_text(aes(label = if_else(filters %in% c('academic', 'geographic', 'demographic') & univ_label == 'Research', str_c(str_to_title(filters), ' filters'), ''), y = 0), hjust = 0, vjust = 1, size = 2, fontface = 'bold') +
  geom_text(aes(label = if_else(!filters %in% c('academic', 'geographic', 'demographic'), str_c(round(pct * 100), '%'), '')), hjust = -0.1, size = 2) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('Number of orders') +
  facet_wrap(~ factor(univ_label, levels = c('Research', 'MA/doctoral'))) +
  coord_flip()
```


## Figure 9: GPA filter used by research vs. ma/doctoral

```{r, fig.height = 3}
orders_gpa_total <- orders_gpa %>% 
  group_by(univ_type) %>% 
  summarise(
    num_orders = sum(n_low)
  ) %>% 
  pull(num_orders, univ_type)

orders_gpa %>% 
  mutate(
    univ_label = recode_factor(
      univ_type,
      'research' = str_c('Research (N=', orders_gpa_total[['research']], ')'),
      'regional' = str_c('MA/doctoral (N=', orders_gpa_total[['regional']], ')')
    )
  ) %>% 
  ggplot(aes(x = factor(gpa_low, levels = c('A+', 'A', 'A-', 'B+', 'B', 'B-', 'C+', 'C', 'C-')), y = pct_low, fill = univ_label)) +
  geom_bar(position = 'dodge', stat = 'identity') +
  geom_text(aes(label = n_low), hjust = 0.5, vjust = 0, size = 2, position = position_dodge(0.9)) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_fill_manual(values = color_palette, name = 'University type') +
  xlab('') + ylab('% of orders in university type')
```


## Figure 10: SAT filter used by research vs. ma/doctoral

```{r, fig.height = 4}
orders_sat_total <- orders_sat %>% 
  filter(test_range == 'sat_min') %>% 
  group_by(univ_type) %>% 
  summarise(
    num_orders = sum(num)
  ) %>% 
  pull(num_orders, univ_type)

orders_sat %>% 
  mutate(
    range_label = recode_factor(
      test_range,
      'sat_min' = 'Minimum',
      'sat_max' = 'Maximum'
    ),
    univ_label = recode_factor(
      univ_type,
      'research' = str_c('Research (N=', orders_sat_total[['research']], ')'),
      'regional' = str_c('MA/doctoral (N=', orders_sat_total[['regional']], ')')
    )
  ) %>% 
  ggplot(aes(x = brks, y = pct, fill = univ_label)) +
  geom_bar(position = position_dodge2(reverse = T, padding = 0), stat = 'identity') +
  geom_text(aes(label = num), hjust = -0.1, size = 2, position = position_dodge2(0.9, reverse = T, padding = 0)) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.1)), labels = label_percent(accuracy = 1)) +
  scale_fill_manual(values = color_palette, name = 'University type') +
  xlab('') + ylab('% of orders in university type') +
  facet_wrap(~ range_label) +
  coord_flip()
```


## Figure 11: PSAT filter used by research vs. ma/doctoral

```{r, fig.height = 4}
orders_psat_total <- orders_psat %>% 
  filter(test_range == 'psat_min') %>% 
  group_by(univ_type) %>% 
  summarise(
    num_orders = sum(num)
  ) %>% 
  pull(num_orders, univ_type)

orders_psat %>% 
  mutate(
    range_label = recode_factor(
      test_range,
      'psat_min' = 'Minimum',
      'psat_max' = 'Maximum'
    ),
    univ_label = recode_factor(
      univ_type,
      'research' = str_c('Research (N=', orders_psat_total[['research']], ')'),
      'regional' = str_c('MA/doctoral (N=', orders_psat_total[['regional']], ')')
    )
  ) %>% 
  ggplot(aes(x = brks, y = pct, fill = univ_label)) +
  geom_bar(position = position_dodge2(reverse = T, padding = 0), stat = 'identity') +
  geom_text(aes(label = num), hjust = -0.1, size = 2, position = position_dodge2(0.9, reverse = T, padding = 0)) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.1)), labels = label_percent(accuracy = 1)) +
  scale_fill_manual(values = color_palette, name = 'University type') +
  xlab('') + ylab('% of orders in university type') +
  facet_wrap(~ range_label) +
  coord_flip()
```


\newpage
## Figure 12: State filter used by research universities

```{r, fig.height = 2.8}
# Out-of-state
plot_usmap(
  regions = 'states',
  data = orders_state_research %>% filter(locale == 'outofstate'),
  values = 'frequency'
) +
  scale_fill_gradient2(
    low = 'gray',
    mid = 'lightgray',
    high = 'cyan', 
    midpoint = 35,
    name = 'Number of orders',
    label = label_number(accuracy = 1),
    limits = c(0, 120)
  ) + 
  labs(title = 'State filter used by research universities (out-of-state)') +
  theme(
    text = element_text(size = 7), 
    plot.title = element_text(color = '#444444', hjust = 0.5, face = 'bold'), 
    legend.title = element_text(face = 'bold'),
    legend.position = 'right'
  )
```

```{r, fig.height = 2.8}
# In-state
plot_usmap(
  regions = 'states',
  data = orders_state_research %>% filter(locale == 'instate'),
  values = 'frequency'
) +
  scale_fill_continuous(
    low = 'lightgray',
    high = 'cyan', 
    name = 'Number of orders',
    label = label_number(accuracy = 1),
    limits = c(0, 10)
  ) + 
  labs(title = 'State filter used by research universities (in-state)') +
  theme(
    text = element_text(size = 7), 
    plot.title = element_text(color = '#444444', hjust = 0.5, face = 'bold'), 
    legend.title = element_text(face = 'bold'),
    legend.position = 'right'
  )
```


## Figure 13: TBD

```{r}

```


## Figure 14: Number of prospects purchased by university type and location

```{r, fig.height = 3}
rq2_counts %>% 
  mutate(
    region = recode_factor(
      region,
      'international' = 'International',
      'domestic' = 'Domesetic'
    ),
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'regional_outofstate' = 'MA/doctoral (out-of-state)',
      'regional_instate' = 'MA/doctoral (in-state)',
      'research_outofstate' = 'Research (out-of-state)',
      'research_instate' = 'Research (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = n, fill = region)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  geom_text(mapping = aes(label = if_else(univ_type == 'regional' & locale == 'outofstate', '', if_else(n < 1000000, str_c(round(n / 1000), 'K'), str_c(round(n / 1000000, 1), 'M')))), size = 2, position = position_stack(vjust = 0.5)) +
  geom_text(mapping = aes(label = if_else(univ_type == 'regional' & region == 'Domesetic' & locale == 'outofstate', str_c(round(n / 1000), 'K (300 international)'), ''), x = 30e3), hjust = 0, size = 2) +
  scale_fill_manual(values = color_palette, name = '') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), labels = function(n) if_else(n < 1000000, str_c(n / 1000, 'K'), str_c(n / 1000000, 'M'))) +
  xlab('Number of prospects') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


## Figure 15: Racial composition of prospects purchased by research universities

```{r, fig.height = 2}
rq2_race %>% 
  filter(univ_type == 'research', region != 'international') %>% 
  mutate(
    stu_race_cb = if_else(is.na(stu_race_cb), 999, unclass(stu_race_cb)),
    race = recode_factor(
      stu_race_cb,
      `999` = 'Missing',
      `0` = 'No response',
      `10` = 'Other',
      `12` = 'Multiracial',
      `1` = 'AI/AN',
      `8` = 'NH/PI',
      `4` = 'Latinx',
      `3` = 'Black',
      `2` = 'Asian',
      `9` = 'White'
    ),
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'research_outofstate' = 'Research (out-of-state)',
      'research_instate' = 'Research (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = pct, fill = race)) + 
  geom_bar(position = 'stack', stat = 'identity', alpha = 0.4) + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = 'BrBG', direction = -1, name = 'Race') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04))) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('Percent of prospects') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


## Figure 16: Median household income of prospects purchased by research universities

```{r, fig.height = 1.5}
rq2_income %>% 
  filter(univ_type == 'research', region != 'international') %>% 
  mutate(
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'research_outofstate' = 'Research (out-of-state)',
      'research_instate' = 'Research (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = income_2564)) + 
  geom_bar(stat = 'identity', fill = color_palette[[2]]) + 
  geom_text(mapping = aes(label = str_c('$', round(income_2564 / 1000), 'K')), size = 2, hjust = -0.1) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), limits = c(0, 115000)) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


## Figure 17: Locale of prospects purchased by research universities

```{r, fig.height = 2}
rq2_locale %>% 
  filter(univ_type == 'research', region != 'international') %>% 
  mutate(
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'research_outofstate' = 'Research (out-of-state)',
      'research_instate' = 'Research (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = pct, fill = locale_text)) + 
  geom_bar(position = 'stack', stat = 'identity', alpha = 0.4) + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = 'BrBG', direction = -1, name = 'Locale') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04))) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('Percent of prospects') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


## Figure 18: Racial composition of prospects purchased by ma/doctoral universities

```{r, fig.height = 2}
rq2_race %>% 
  filter(locale == 'instate', region != 'international') %>% 
  mutate(
    stu_race_cb = if_else(is.na(stu_race_cb), 999, unclass(stu_race_cb)),
    race = recode_factor(
      stu_race_cb,
      `999` = 'Missing',
      `0` = 'No response',
      `10` = 'Other',
      `12` = 'Multiracial',
      `1` = 'AI/AN',
      `8` = 'NH/PI',
      `4` = 'Latinx',
      `3` = 'Black',
      `2` = 'Asian',
      `9` = 'White'
    ),
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'research_instate' = 'Research (in-state)',
      'regional_instate' = 'MA/doctoral (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = pct, fill = race)) + 
  geom_bar(position = 'stack', stat = 'identity', alpha = 0.4) + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = 'BrBG', direction = -1, name = 'Race') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04))) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('Percent of prospects') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


## Figure 19: Median household income of prospects purchased by ma/doctoral universities

```{r, fig.height = 2}
rq2_income %>% 
  filter(locale == 'instate', region != 'international') %>% 
  mutate(
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'research_instate' = 'Research (in-state)',
      'regional_instate' = 'MA/doctoral (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = income_2564)) + 
  geom_bar(stat = 'identity', fill = color_palette[[2]]) + 
  geom_text(mapping = aes(label = str_c('$', round(income_2564 / 1000), 'K')), size = 2, hjust = -0.1) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04)), limits = c(0, 105000)) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


## Figure 20: Locale of prospects purchased by ma/doctoral universities

```{r, fig.height = 2}
rq2_locale %>% 
  filter(locale == 'instate', region != 'international') %>% 
  mutate(
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'research_instate' = 'Research (in-state)',
      'regional_instate' = 'MA/doctoral (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = pct, fill = locale_text)) + 
  geom_bar(position = 'stack', stat = 'identity', alpha = 0.4) + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = 'BrBG', direction = -1, name = 'Locale') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04))) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('Percent of prospects') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


## Figure A1: School type of prospects purchased by research vs. ma/doctoral universities

```{r, fig.height = 2.5}
rq2_school %>% 
  filter(univ_type == 'research' | locale == 'instate', region != 'international') %>% 
  mutate(
    control = if_else(is.na(hs_school_control), 'unknown', hs_school_control),
    control = recode_factor(
      control,
      'unknown' = 'Missing',
      'private' = 'Private HS',
      'public' = 'Public HS'
    ),
    count_type = str_c(univ_type, locale, sep = '_'),
    count_type = recode_factor(
      count_type,
      'regional_instate' = 'MA/doctoral (in-state)',
      'research_outofstate' = 'Research (out-of-state)',
      'research_instate' = 'Research (in-state)'
    )
  ) %>% 
  ggplot(aes(y = count_type, x = pct, fill = control)) + 
  geom_bar(position = 'stack', stat = 'identity', alpha = 0.4) + 
  geom_text(mapping = aes(label = if_else(pct > 0.02, as.character(round(pct * 100)), '')), size = 2, position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = 'PuBuGn', direction = 1, name = 'School type') +
  scale_x_continuous(expand = expansion(mult = c(0, 0.04))) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.1))) +
  xlab('Percent of prospects') + ylab('') +
  guides(fill = guide_legend(reverse = T)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = '#444444', size = 7, face = 'bold', hjust = 1)
  )
```


## Table 7: Filter combos used in order purchases by research vs. ma/doctoral

```{r}
bind_cols(
  orders_filters_combo %>% 
    filter(univ_type == 'research') %>% 
    select(filter_combos, n, pct) %>% 
    head(10) %>%
    mutate(pct = str_c(round(pct * 100), '%')),
  orders_filters_combo %>% 
    filter(univ_type == 'regional') %>% 
    select(filter_combos, n, pct) %>% 
    head(10) %>%
    mutate(pct = str_c(round(pct * 100), '%'))
) %>% 
  kable(booktabs = T, col.names = rep(c('Filters', 'Count', 'Percent'), 2), align = rep(c('l', 'c', 'c'), 2)) %>%
  add_header_above(c('Research' = 3, 'MA/doctoral' = 3), bold = T) %>% 
  row_spec(0, bold = T) %>%
  kable_styling(position = 'center', latex_options = c('hold_position', 'scale_down'))
```


## Table 8: Prospect characteristics by filter used

```{r}
rq3 %>% 
  mutate_if(is.numeric, ~if_else(. < 1, . * 100, .)) %>% 
  mutate_if(is.numeric, round, 0) %>% 
  mutate_if(is.numeric, ~if_else(row_number() == 18, str_c('$', round(. / 1000), 'K'), prettyNum(., big.mark = ','))) %>% 
  mutate(
    row_subj = recode_factor(
      row_subj,
      'n' = 'Total count',
      'pct_instate' = '% In-state',
      'pct_outofstate' = '% Out-of-state',
      'pct_white' = '% White',
      'pct_asian' = '% Asian',
      'pct_black' = '% Black',
      'pct_latinx' = '% Latinx',
      'pct_nhpi' = '% NH/PI',
      'pct_aian' = '% AI/AN',
      'pct_multiracial' = '% Multiracial',
      'pct_raceother' = '% Other',
      'pct_noresponse' = '% No response',
      'pct_racemissing' = '% Missing',
      'pct_male' = '% Male',
      'pct_female' = '% Female',
      'pct_genderother' = '% Other ',
      'pct_gendermissing' = '% Missing ',
      'med_income' = 'Median income',
      'pct_city' = '% City',
      'pct_suburban' = '% Suburban',
      'pct_fringe' = '% Rural - Fringe',
      'pct_distant' = '% Rural - Distant',
      'pct_remote' = '% Rural - Remote',
      'pct_localemissing' = '% Missing  '
    )
  ) %>%
  arrange(row_subj) %>% 
  kable(booktabs = T, col.names = c('', 'All domestic', 'GPA', 'PSAT', 'SAT', 'HS rank', 'AP score', 'Zip code', 'State', 'Geomarket', 'Segment', 'CBSA', 'Race', 'Gender'), align = c('l', rep('r', 13))) %>% 
  add_header_above(c(' ' = 2, 'Academic' = 5, 'Geographic' = 5, 'Demographic' = 2), bold = T) %>% 
  group_rows('Location', 2, 3) %>%
  group_rows('Race/ethnicity', 4, 13) %>%
  group_rows('Gender', 14, 17) %>%
  group_rows('Household income', 18, 18) %>%
  group_rows('Locale', 19, 24) %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = c('hold_position', 'scale_down'))
```
