---
title: "Texas A & M University - Texarkana"
urlcolor: blue
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    fig_caption: true
    highlight: tango
    theme: default
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(tibble.width = Inf, width = 10000, scipen = 999)

library(tidyverse)
library(viridis)
library(leaflet)
library(rgdal)
library(formattable)

load(url('https://github.com/mpatricia01/public_requests_eda/raw/main/data/224545_data.RData'))

remove_NA_cols <- function(data_df) {
  data_df[!sapply(data_df, function(x) all(is.na(x)))]
}

sort_rows <- function(data_df) {
  data_df %>% arrange(date_start, hs_grad_class, sat_score_min, sat_score_max, psat_score_min, psat_score_max)
}

primary_orders <- orders_df %>% filter(market == 'primary') %>% remove_NA_cols() %>% sort_rows()
secondary_orders <- orders_df %>% filter(market == 'secondary') %>% remove_NA_cols() %>% sort_rows()
houston_orders <- orders_df %>% filter(market == 'houston') %>% remove_NA_cols() %>% sort_rows()
dallas_orders <- orders_df %>% filter(market == 'dallas') %>% remove_NA_cols()

zip_data <- read_csv(url('https://raw.githubusercontent.com/cyouh95/third-way-report/master/assets/data/zip_to_state.csv')) %>% 
  mutate(pop_poc_pct = (pop_black + pop_hispanic + pop_amerindian) / pop_total * 100)
zip_shp <- readOGR(file.path('..', 'data', 'cb_2018_us_zcta510_500k', 'cb_2018_us_zcta510_500k.shp'))
```

<style>
pre code, pre, code {
  white-space: pre !important;
  overflow-x: auto !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
</style>


# Primary market purchases (38 orders)

```{r}
primary_orders
```

Generally, on each purchase date, they make **orders by SAT or PSAT score ranges** (**SAT** if purchasing _seniors_ or _juniors_, **PSAT** if purchasing _juniors_ or _sophomores_):

- **Senior purchases** by **SAT** filter
  - **Sep 2017** (_grad class 2018_): 4 groups (A, B1, C, D)
  - **Sep 2018** (_grad class 2019_): 4 groups (A, B2, C, D)
  - **Sep 2019** (_grad class 2020_): 1 group (B3)
- **Junior purchases** by **SAT** or **PSAT** filter
  - **Sep 2017** (_grad class 2019_): 3 groups (C1, D, E) - **PSAT**
  - **Feb 2018** (_grad class 2019_): 3 groups (B, C1, D) - **PSAT**
  - **Sep 2018** (_grad class 2020_): 3 groups (C2, D, E) - **PSAT**
  - **Feb 2019** (_grad class 2020_): 4 groups (B, C2, D, E) - **PSAT**
  - **Sep 2019** (_grad class 2021_): 1 group (B3) - **SAT**
  - **Apr 2020** (_grad class 2021_): 1 group (B3) - **SAT**
- **Sophomore purchases** by **PSAT** filter
  - **Sep 2017** (_grad class 2020_): 2 groups (B, E)
  - **Feb 2018** (_grad class 2020_): 4 groups (A, B, C1, D)
  - **Sep 2018** (_grad class 2021_): 3 groups (C2, D, E)
  - **Feb 2019** (_grad class 2021_): 4 groups (B, C2, D, E)
  - **Sep 2019** (_grad class 2022_): 1 group (C3)

```{r}
distinct(primary_orders %>% select(sat_score_min, sat_score_max, psat_score_min, psat_score_max)) %>% 
  arrange(sat_score_min, sat_score_max, psat_score_min, psat_score_max) %>%
  add_column(score_range = c('A', 'B1', 'B2', 'B3', 'C', 'D', 'A', 'B', 'C1', 'C2', 'C3', 'D', 'E'), .before = 1)
```

**Common filters**:

- **GPA**: A+ to C+

**General observations**:

- Only purchase **seniors** during the Fall, and **juniors** and **sophomores** both Fall & Spring
  - For **juniors** and **sophomores**, Spring purchases generally have lower test score ranges than Fall purchases
  - Over the years, there is a slight increase in some test score ranges for all grad classes
- Usually making multiple orders at a time w/ different test score ranges, but starting in 2019, only made 1 order covering broader score range for all grad classes
  - Also, started using SAT instead of PSAT filter for **sophomores**


# Secondary market purchases (32 orders)

```{r}
secondary_orders
```


# Houston market purchases (20 orders)

```{r}
houston_orders
```


# Dallas market purchases (1 order)

```{r}
dallas_orders
```


# Order summaries

```{r}
# Cost per student went from $0.43 to $0.45 (starting Fall 2018) to $0.47 (starting Fall 2019)
orders_rate_df <- orders_df %>% select(order_num, date_start, order_cost, num_students) %>% mutate(rate = round(order_cost / num_students, 2)) %>% arrange(rate, date_start)
orders_rate_df

orders_rate_df %>%
  ggplot(aes(x = as.character(order_num), y = num_students, fill = as.character(date_start))) + 
  geom_bar(position = 'stack', stat = 'identity') +
  scale_fill_viridis(discrete = T) +
  labs(x = '') +
  theme(
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank()
  ) +
  facet_grid(~rate, scales = 'free_x', space = 'free_x')

orders_rate_df %>%
  ggplot(aes(x = as.character(order_num), y = order_cost, fill = as.character(date_start))) + 
  geom_bar(position = 'stack', stat = 'identity') +
  scale_fill_viridis(discrete = T) +
  labs(x = '') +
  theme(
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank()
  ) +
  facet_grid(~rate, scales = 'free_x', space = 'free_x')
```


# Maps

```{r}
# Create var for race breaks
zip_data$race_brks_nonwhiteasian <- cut(zip_data$pop_poc_pct, 
                                        breaks = c(-1, 20, 40, 60, 80, 90, 101), 
                                        labels = c('0-19%', '20-39%', '40-59%', 
                                                   '60-79%', '80-89%', '90-100%'))

# Create var for income breaks
zip_data$inc_brks <- cut(zip_data$median_household_income, 
                         breaks = c(-1, 50000, 75000, 100000, 150000, 200000, 10000000), 
                         labels = c('<$50k', '$50k-74k', '$75k-99k', 
                                    '$100k-149k', '$150k-199k', '$200k+'))

# Create var for first 3 digits of zip code
zip_data$zip_code_3 <- zip_data$zip_code %>% str_sub(end = 3)

# Merge zip data to shapes
zip_shp <- merge(zip_shp, zip_data, by.x = 'ZCTA5CE10', by.y = 'zip_code', all.x = T)


# Create shared color scale functions
color_income <- colorFactor('YlGnBu', zip_shp$inc_brks)
color_race <- colorFactor('YlGnBu', zip_shp$race_brks_nonwhiteasian)
color_pop <- colorNumeric('YlGnBu', zip_shp$pop_total, n = 5)


# Plot purchased zip codes
plot_map <- function(map_data) {
  # Create popups
  pop_zip <- paste0('<b>', zip_shp_purchased$zip_name, '</b><br>',
                    'Total Population: ', format(zip_shp_purchased$pop_total, big.mark = ',')) %>% lapply(htmltools::HTML)
  
  income_zip <- paste0('<b>', zip_shp_purchased$zip_name, '</b><br>',
                       'Median Household Income: ', currency(zip_shp_purchased$median_household_income, digits = 0L)) %>% lapply(htmltools::HTML)
  
  race_zip <- paste0('<b>', zip_shp_purchased$zip_name, '</b><br>',
                     '% Population of Color: ', sprintf('%.1f', zip_shp_purchased$pop_poc_pct)) %>% lapply(htmltools::HTML)
  
  leaflet(data = map_data) %>%
    addProviderTiles(providers$CartoDB.Positron) %>%
    
    addMiniMap(tiles = providers$CartoDB.Positron,
               toggleDisplay = TRUE) %>% 
    
    addPolygons(stroke = F, fillOpacity = 0.1, smoothFactor = 0.2, color = 'gray', group = 'Purchased') %>% 
    addPolygons(stroke = F, fillOpacity = 0.8, smoothFactor = 0.2, color = ~color_pop(pop_total), label = pop_zip, group = 'By Population') %>%
    addPolygons(stroke = F, fillOpacity = 0.8, smoothFactor = 0.2, color = ~color_income(inc_brks), label = income_zip, group = 'By Median Household Income') %>%
    addPolygons(stroke = F, fillOpacity = 0.8, smoothFactor = 0.2, color = ~color_race(race_brks_nonwhiteasian), label = race_zip, group = 'By Race/Ethnicity') %>% 
    
    # add legends
    addLegend(position = 'topright', pal = color_pop, values = ~pop_total,
              title = 'Population',
              className = 'info legend legend-pop',
              na.label = 'NA',
              opacity = 1) %>%
    
    addLegend(position = 'topright', pal = color_income, values = ~inc_brks,
              title = 'Median Household Income',
              className = 'info legend legend-income',
              na.label = 'NA',
              opacity = 1) %>%
    
    addLegend(position = 'topright', pal = color_race, values = ~race_brks_nonwhiteasian,
              title = 'Black, Latinx, and <br>Native American Population',
              className = 'info legend legend-race',
              na.label = 'NA',
              opacity = 1) %>%
    
    # add options
    addLayersControl(
      position = c('bottomleft'),
      baseGroups = c('Purchased', 'By Population', 'By Median Household Income', 'By Race/Ethnicity'),
      options = layersControlOptions(collapsed = FALSE)
    ) %>%
    
    htmlwidgets::onRender("
          function(el, x) {
              var myMap = this;
              $('.legend').css('display', 'none');
              
              myMap.on('baselayerchange', function(e) {
                  $('.legend').css('display', 'none');
                  switch(e.name) {
                      case 'By Population':
                          $('.legend-pop').css('display', 'inherit');
                          break;
                      case 'By Median Household Income':
                          $('.legend-income').css('display', 'inherit');
                          break;
                      case 'By Race/Ethnicity':
                          $('.legend-race').css('display', 'inherit');
                          break;
                  }
                  e.layer.bringToBack();
              });
      }")
}

# View different sets of zip codes purchased
zip_purchased <- str_split(unique(orders_df$zip_code) %>% na.omit() %>% as.character(), '\\|')

# Primary
zip_shp_purchased <- subset(zip_shp, zip_code_3 %in% zip_purchased[[2]])
plot_map(zip_shp_purchased)

# Secondary (1, 4, 5)
zip_shp_purchased <- subset(zip_shp, zip_code_3 %in% zip_purchased[[1]])
plot_map(zip_shp_purchased)

# Houston
zip_shp_purchased <- subset(zip_shp, zip_code_3 %in% zip_purchased[[3]])
plot_map(zip_shp_purchased)
```
